A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change

### 2018-11-29 paper1_tKdV/tKdV_theory.tex
NM

\affil[a]{Department of Mathematics and Center for Atmosphere
and Ocean Science, Courant Institute of Mathematical Sciences, New
York University, New York, NY 10012}
\affil[b]{Department of Mathematics and Geophysical Fluid
Dynamics Institute, Florida State University, Tallahassee, FL}

@@

Consider a spatial discretization with $J=2\Lambda+1$ grid points, so that the smallest
resolved scale is comparable to the characteristic wavelength
\[
\Delta x=\frac{2\pi M\lambda_{c}}{J}\lesssim\lambda_{c}\:\Rightarrow\:M=\frac{J}{2\pi}\sim5,\;J=32.
\]
In the practical numerical simulations, we select $M=5$ and
let $\gamma$ vary in the range $\left[0.5,1\right]$. Using reference values from the experiments, 
\cite{bolles2018anomalous}, we find $\epsilon\in\left[0.0024,0.024\right], \delta\sim0.22$,
and $D_{0}$ changes from 1 to 0.24 at the ADC.
These values yield the following the following ranges for the model parameters: $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
These are the values we will test in the direct numerical simulations.
See details about the derivation from scale analysis in \textcolor{blue}{\emph{SI Appendix, A}}.

@@

The Gibbs measures $\mathcal{G}_{\theta}^{\pm}$ are defined based
on the different inverse temperatures $\theta^{\pm}$ on the two sides
of the solutions
\[
\begin{aligned}\mu_{t}^{-}\left(u_{\Lambda}^{-};D_{-}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}-}}=u_{0},\;t<T_{\mathrm{ADC}};\\
\mu_{t}^{+}\left(u_{\Lambda}^{+};D_{+}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}+}}=u_{0},\;t>T_{\mathrm{ADC}},
\end{aligned}
\]
where $u_{0}$ represents the deterministic matching condition between
the incoming and outgoing waves. The two distributions, $\mu_{t}^{-},\mu_{t}^{+}$
should be matched at $T_{\mathrm{ADC}}$ giving
\[
\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
\]
In matching the flow statistics before and after the abrupt depth
change, we first use the conservation of the deterministic Hamiltonian
$H_{\Lambda}^{+}$ after the depth change. Then assuming ergodicity
\cite{abramov2003hamiltonian,majda2006nonlinear}, the statistical
expectation for the Hamiltonian $\left\langle H_{\Lambda}^{+}\right\rangle $
is conserved in time after the depth change at $t=T_{\mathrm{ADC}}$
and should stay in the same value as the system approaches equilibrium
as $t\rightarrow\infty$. The final statistical matching condition
to get the outgoing flow statistics with parameter $\theta^{+}$ can
be found by
\begin{equation}
\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{+}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}},\label{eq:matching}
\end{equation}
with the outgoing flow Hamiltonian $H_{\Lambda}^{+}$ and the Gibbs
measures $\mathcal{G}_{\theta}^{\pm}$ before and after the abrupt
depth change.

>>>>

\affil[a]{Department of Mathematics and Center for Atmosphere
and Ocean Science, Courant Institute of Mathematical Sciences, New
York University, New York, NY 10012}
\affil[b]{Department of Mathematics and Geophysical Fluid
Dynamics Institute, Florida State University, Tallahassee, FL 32306}

@@

Consider a spatial discretization with $J=2\Lambda+1$ grid points, so that the smallest
resolved scale is comparable to the characteristic wavelength
\[
\Delta x=\frac{2\pi M\lambda_{c}}{J}\lesssim\lambda_{c}\:\Rightarrow\:M=\frac{J}{2\pi}\sim5,\;J=32.
\]
In the practical numerical simulations, we select $M=5$ and
let $\gamma$ vary in the range $\left[0.5,1\right]$. Using reference values from the experiments, 
\cite{bolles2018anomalous}, we find $\epsilon\in\left[0.0024,0.024\right], \delta\sim0.22$,
and $D_{0}$ changes from 1 to 0.24 at the ADC.
These values yield the following ranges for the model parameters: $L_{0}\in\left[2,6\right]$ and $E_{0}\in\left[50,200\right]$.
These are the values we will test in the direct numerical simulations.
See details about the derivation from scale analysis in \textcolor{blue}{\emph{SI Appendix, A}}.

@@

The Gibbs measures $\mathcal{G}_{\theta}^{\pm}$ are defined based
on the different inverse temperatures $\theta^{\pm}$ on the two sides
of the solutions
\[
\begin{aligned}\mu_{t}^{-}\left(u_{\Lambda}^{-};D_{-}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}-}}=u_{0},\;t<T_{\mathrm{ADC}};\\
\mu_{t}^{+}\left(u_{\Lambda}^{+};D_{+}\right), & \quad u_{\Lambda}\mid_{t=T_{\mathrm{ADC}+}}=u_{0},\;t>T_{\mathrm{ADC}},
\end{aligned}
\]
where $u_{0}$ represents the deterministic matching condition between
the incoming and outgoing waves. The two distributions, $\mu_{t}^{-},\mu_{t}^{+}$
should be matched at $T_{\mathrm{ADC}}$ giving
\[
\mu_{t=T_{\mathrm{ADC}}}^{-}\left(u_{\Lambda}\right)=\mu_{t=T_{\mathrm{ADC}}}^{+}\left(u_{\Lambda}\right).
\]
In matching the flow statistics before and after the abrupt depth
change, we first use the conservation of the deterministic Hamiltonian
$H_{\Lambda}^{+}$ after the depth change. Then assuming ergodicity
\cite{abramov2003hamiltonian,majda2006nonlinear}, the statistical
expectation for the Hamiltonian $\left\langle H_{\Lambda}^{+}\right\rangle $
is conserved in time after the depth change at $t=T_{\mathrm{ADC}}$
and should remain at this value as the system approaches equilibrium
as $t\rightarrow\infty$. The final statistical matching condition
to get the outgoing flow statistics with parameter $\theta^{+}$ can
be found by
\begin{equation}
\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{+}}=\left\langle H_{\Lambda}^{+}\right\rangle _{\mathcal{G}_{\theta}^{-}},\label{eq:matching}
\end{equation}
with the outgoing flow Hamiltonian $H_{\Lambda}^{+}$ and the Gibbs
measures $\mathcal{G}_{\theta}^{\pm}$ before and after the abrupt
depth change.
